---
title: "Homework_1"
author: "Shadman Amin & Lily Heidger"
format: 
  html:
    code-fold: show
    toc: true
    embed-resources: true
editor: visual
theme: lumen
execute:
  echo: true
  message: false
  warning: false
---

# Homework 1

```{r, setup}
library(dplyr)
library(here)
library(readr)
library(tidyverse)

algae <- read_table("algaeBloom.txt", col_names= c('season', 'size', 'speed', 'mxPH', 'mnO2', 'Cl', 'NO3', 'NH4', 'oPO4','PO4','Chla','a1','a2','a3','a4','a5','a6','a7'),
                     na="XXXXXXX")
glimpse(algae)
```

## 1 Descriptive Summary Statistics

### a)

```{r}
algae %>% group_by(season) %>% summarise(length(size))
```

### b)

Yes, there are 33 missing values. For most of the chemicals (Cl, NO3, NH4, oPO4, PO4, Chla), the variances are significantly larger than the means, often several times or several orders of magnitude larger. However, for mxPH and mnO2, the means are significantly larger than the variance. Yet for all of the chemicals, the means and variances are significantly different.

```{r}
options(scipen=999)

chemicals <- algae %>% select(mxPH, mnO2, Cl, NO3, NH4, oPO4, PO4, Chla)
chems <- c("mxPH", "mnO2", "Cl", "NO3", "NH4", "oPO4", "PO4", "Chla")
means <- c(colMeans(chemicals, na.rm = TRUE))
means_df <- data.frame(means, chems)

vars <- c(sapply(chemicals, var, na.rm = TRUE))
vars_df <- data.frame(vars, chems)

chems <- left_join(vars_df, means_df) %>% select(chems, means, vars)

kableExtra::kable(chems)
```

### c)

The median and MAD are much more similar for almost all of the chemicals compared to the mean and (confuseddddddd)

Lower variance- median and mad more similar

```{r}
medians <- sapply(chemicals, median, na.rm = TRUE)
medians_df <- data.frame(medians, chems)

mxPH_mad <- mad(chemicals$mxPH, na.rm = TRUE)
mnO2_mad <- mad(chemicals$mnO2, na.rm = TRUE)
Cl_mad <- mad(chemicals$Cl, na.rm = TRUE)
NO3_mad <- mad(chemicals$NO3, na.rm = TRUE)
NH4_mad <- mad(chemicals$NH4, na.rm = TRUE)
oPO4_mad <- mad(chemicals$oPO4, na.rm = TRUE)
PO4_mad <- mad(chemicals$PO4, na.rm = TRUE)
Chla_mad <- mad(chemicals$Chla, na.rm = TRUE)

mad <- data.frame(chems, mads = c(mxPH_mad,mnO2_mad,Cl_mad,NO3_mad,NH4_mad,oPO4_mad,PO4_mad,Chla_mad)) 

all <- right_join(mad, medians_df)
```

## 2.

### a)

The distribution is skewed left

```{r}
ggplot(chemicals, aes(mxPH, after_stat(density)))+
  geom_histogram() +
  ggtitle("Histogram of mxPH")

```

```{r}
ggplot(algae, aes(mxPH))+
  geom_histogram(aes(y = stat(density))) +
  geom_density() +
  geom_rug() +
  ggtitle("Histogram of mxPH")
```

### c)

Medium speeds of a4 seem to have a larger variance than the other two speeds, with low speeds having the lowest variance. All three speeds seem to have similar lower quartiles and medians, both of which are close to zero.

```{r}
a4_df <- algae %>% select(speed, a4) %>% group_by(speed)

ggplot(a4_df, aes(a4, speed)) + geom_boxplot() +
  ggtitle("A conditioned Boxplot of Algal a4")
```

## 3.

### a)

There are 33 missing values. Season, size, and speed have 0 missing values; mxPH has 1, mnO2, NO3, NH4, oPO4, and PO4 have 2; and Cl has 10 missing values.

```{r}
na_count <-sapply(algae, function(y) sum(length(which(is.na(y)))))

data.frame(na_count)
sum(na_count)
```

### b)

```{r}
nas <- is.na(algae)
algae.del <- algae %>% filter(is.na())
```

## 4. a)

## 5.

Bias()

## 7.

### 1)

Let $x = \begin{bmatrix} x_1\\ x_2\\ .. \\ x_p\\ \end{bmatrix}$ and $y = \begin{bmatrix} y_1\\ y_2\\ .. \\ y_p\\ \end{bmatrix}$. Therefore, $x-y = \begin{bmatrix} x_1 - y_1\\ x_2 - y_2\\ .. \\ x_p - y_p\\ \end{bmatrix}$.

$\|x-y\|_2$ = $\sqrt{\sum_{i=1}^{p} \left ( x_i -y_i\right )^2}$

Law of Positivity: Since this equation is the square root of a sum of non-negative numbers, by definition $\|x-y\|_2 \geq 0$ is true, it is always positive

Law of symmetry: $\|x-y\|_2$ = 0 when all $x_i = y_i$ is true, which is the case only when $x=y$.

$d(x,y) = \|x-y\|_2$ = $\sqrt{\sum_{i=1}^{p} \left ( x_i -y_i\right )^2}$

$d(y,x) = \|y-x\|_2$ = $\sqrt{\sum_{i=1}^{p} \left ( y_i -x_i\right )^2}$

For every $x_i, y_i$ pair, $\left | x_i - y_i \right | = \left | y_i - x_i \right |$.

Hence, $\sqrt{\sum_{i=1}^{p} \left ( x_i -y_i\right )^2} = \sqrt{\sum_{i=1}^{p} \left ( y_i -x_i\right )^2}$.

It shows, $d(x,y) = d(y,x)$

Triangle inequality:

Let,

$z = \begin{bmatrix} z_1\\ z_2\\ .. \\ z_p\\ \end{bmatrix}$

Hence,

$\|z\|_2$ = $\sqrt{\sum_{i=1}^{p} z_1^2 +z_2^2 + z_3^2 + ...+z_p^2}$

We need to show

$||x-z||_2$ $\leq$ $||x-y||_2 + ||y-z||_2$

Taking $a = x-y$ and $b = y-z$ , we can use Schwarz inequality to show that $||a+b||_2 \leq ||a||_2 + ||b||_2$ which gives us

$d(x,z) \leq d(x,y) + d(y,z)$ and hence follows the law of triangle inequality

### 2)

By definition, $||x||_{\infty}$ = $max_i (|x|_i)$

Therefore, $||x-y||_{\infty}$ = $max_i (|x_i-y_i|)$
